Simple simulations for a simplified version of web-boost-2

### 2020-07-03 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
mr.ash2s lasso beta init

plot_pip <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

@@

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

@@

  fig1 <- plot_pip(dt.gene)
  fig2 <- plot_beta(dt.gene)
  fig3 <- plot_p(dt.gene)
  fig4 <- plot_pip(dt.snp)
  fig5 <- plot_beta(dt.snp)
  fig6 <- plot_p(dt.snp)

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
    xaxis = list(title = "Genomic position"))

@@

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
```

@@

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)

plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

MR.ASH results (start from SNP):
```{r sim 5 snp, out.width = "98%", out.height="95%"}
tag <- "20200616-5"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)

plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

@@

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)

plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

MR.ASH results (start from SNP):
```{r sim 6 snp, out.width = "98%", out.height="95%"}
tag <- "20200616-6"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)

plot_all(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

>>>>

plot_pip <- function(dt, ld = F){

  if (isTRUE(ld)) {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
                   mode = 'markers', color = ~ max_r2, 
                   text = ~ paste("Name: ", name, 
                                  "\n", ifcausal,
                                  "\nmax R^2: ", round(max_r2, digits = 2), 
                                  "\nMax pair: ", max_pair))
  } else {
    pal <- c("salmon", "darkgreen")
    pal <- setNames(pal, c("Causal", "Non causal"))

    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
  }

@@

plot_susie_pip <- function(dt) {
   pal <- c("salmon", "darkgreen")
   pal <- setNames(pal, c("Causal", "Non causal"))

   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres, ldres){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  ld <- read.table(ldres, header =T)
  a <- merge(a, ld, by = "name", all.x = T)

  su <- read.table(susieres, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

@@

  fig1 <- plot_pip(dt.gene, ld = T)  
  fig2 <- plot_beta(dt.gene) 
  fig3 <- plot_p(dt.gene) 
  fig4 <- plot_susie_pip(dt.gene) 
  fig5 <- plot_pip(dt.snp) 
  fig6 <- plot_beta(dt.snp) 
  fig7 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, nrows=7, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 900, margin = 0.02,
    xaxis = list(title = "Genomic position"))

@@

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
```

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
  susie <- paste0(susiedir, tag, ".", tag2, ".L3-susieres.txt")
  rsq <- paste0(outputdir, tag, ".LD.txt")
  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, susie, rsq))
}
```

@@

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (start from SNP): very similar to start from gene.
```{r sim 5 snp, out.width = "98%", out.height="95%", eval = F}
tag <- "20200616-5"
tag2 <- "snp-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

@@

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (start from SNP): very similar to start from gene
```{r sim 6 snp, out.width = "98%", out.height="95%", eval = F}
tag <- "20200616-6"
tag2 <- "snp-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

## Simulation 7
Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).

Use beta.init = lasso for mr.ash.

MR.ASH results (use lasso for all):
```{r sim 7 lasso, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-7"
tag2 <- "lasso"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (use lasso for just for SNPs):
```{r sim 7 lassosnp, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-7"
tag2 <- "lassoSNP"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

## Simulation 8
Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).

Use beta.init = lasso for mr.ash.

MR.ASH results (use lasso for all):
```{r sim 8 lasso, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-8"
tag2 <- "lasso"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (use lasso for just for SNPs):
```{r sim 8 lassosnp, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-8"
tag2 <- "lassoSNP"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

### 2020-07-03 analysis/simulation_simpleveb-boost2.Rmd
mr.ash2s lasso beta init

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
source("~/causalTWAS/causal-TWAS/code/mr.ash2.R")
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
source("~/causalTWAS/causal-TWAS/code/mr.ash2.R")
```

@@

### init with lasso

```{r run mr.ash2, lasso}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2s(X.gene, X.SNP, y, mr.ash.init = "lasso")

### init with lasso SNP

```{r run mr.ash2, lassosnp}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2s(X.gene, X.SNP, y, mr.ash.init = "lassoX2")
